---
title: "Problem Set 1"
author: "Answer Key"
date: "ECON 480 — Fall 2021"
output:
  html_document:
    df_print: paged
    toc: true 
    toc_depth: 3
    toc_float: true
    code_folding: show
    highlight: tango
# for pdf:
#output:
  #pdf_document:
    #dev: cairo_pdf
    #latex_engine: xelatex
---

<!--CLICK "KNIT" ABOVE TO RENDER TO HTML, PDF, OR WORD OUTPUT
In fact, try knitting right away and see what this produces!

To knit to PDF, you will need to install a LaTeX distribution, which you can do
easily by running the following code in R:

install.packages("tinytex")
tinytex::install_tinytex()
-->

Answers generally go above and beyond what I expect from you. They are meant to show you the correct answer, explain *why* it is correct, and potentially show *several methods* by which you can reach the answer.

```{r message=FALSE, warning=FALSE, include=FALSE}
knitr::opts_chunk$set(echo = T,
                      fig.retina = 3)
```

# The Popularity of Baby Names

Install and load the package `babynames`. Get help for `?babynames` to see what the data includes.

```{r, load-packages}
# install.packages("babynames")

# Note I've “commented” out some of these commands  (with a #) so they do not run when I knit this document
# You should **never** install a package inside a .Rmd document, just do that in R Studio itself
# Of course, you do need to load everything with library() in a .Rmd document!

library(babynames)
library(tidyverse)
# ?babynames()
```

## Question 1 
### Part A
**What are the top 5 boys names for 2017, and what *percent* of overall names is each?**

<!--WRITE YOUR ANSWERS BELOW -->

```{r 1-a}
# save as a new tibble
top_5_boys_2017 <- babynames %>% # take data
  filter(sex=="M", # filter by males
         year==2017) %>% # and for 2007
  arrange(desc(n)) %>% # arrange in largest-to-smallest order of n (number)
  slice(1:5) %>% # optional, look only at first 5 rows; head(., n=5) also works
  mutate(percent = round(prop*100, 2)) # also optional, make a percent variable rounded to 2 decimals

# look at our new tibble
top_5_boys_2017
```

The top 5 names are

```{r}
top_5_boys_2017 %>%
  select(name,percent) %>%
  knitr::kable()
```

Alternatively, you could just write what you found manually into an object like:

```{r}
top_5_boys_2017_alt <- c("Liam", "Noah", "William", "James", "Logan")

top_5_boys_2017_alt
# you could alternatively add a command, 
# %>% pull(name) to the first chunk of code, 
# and it would do the same thing, but we'd want to save it, 
# for example:

top_5_boys_2017_alt <- babynames %>%
  filter(sex=="M",
         year==2017) %>%
  arrange(desc(n)) %>% 
  slice(1:5) %>%
  mutate(percent = round(prop*100, 2)) %>%
  pull(name)
  
top_5_boys_2017_alt
```

### Part B
**What are the top 5 *girls* names for 2017, and what *percent* of overall names is each?**

<!--WRITE YOUR ANSWERS BELOW -->

```{r 1-b}
# save as a new tibble
top_5_girls_2017 <- babynames %>% # take data
  filter(sex=="F", # filter by females
         year==2017) %>% # and for 2007
  arrange(desc(n)) %>% # arrange in largest-to-smallest order of n (number)
  slice(1:5) %>% # optional, look only at first 5 rows; head(., n=5) also works
  mutate(percent = round(prop*100, 2)) # also optional, make a percent variable rounded to 2 decimals

# look at our new tibble
top_5_girls_2017
```

The top 5 names are

```{r}
top_5_girls_2017 %>%
  select(name,percent) %>%
  knitr::kable()
```

Alternatively, you could just write what you found manually into an object like:

```{r}
top_5_girls_2017_alt <- c("Emma", "Olivia", "Ava", "Isabella", "Sophia")
```

## Question 2 
**Make two barplots of these top 5 names, one for each sex. Map `aes`thetics `x` to `name` and `y` to `prop` [or `percent`, if you made that variable, as I did.] and use `geom_col` (since you are declaring a specific `y`, otherwise you could just use `geom_bar()` and just an `x`.)**

<!--WRITE YOUR ANSWERS BELOW -->

```{r 2-boys}
ggplot(data = top_5_boys_2017)+
  aes(x = reorder(name, n), #note this reorders the x variable from small to large n
      y = percent, # you can use prop if you didn't make a percent variable
      fill = name)+ # optional color!
  geom_col()+
  
  # all of the above is sufficient, now I'm just making it pretty
  scale_y_continuous(labels = function(x){paste0(x, "%")}, # add percent signs
                     breaks = seq(from = 0, # make line breaks every 0.25%
                                  to = 1,
                                  by = 0.25),
                     limits = c(0,1), # limit axis to between 0 and 1
                     expand = c(0,0))+ # don't let it go beyond this
  labs(x = "Name",
       y = "Percent of All Babies With Name",
       title = "Most Popular Boys Names Since 1880",
       fill = "Boy's Name",
       caption = "Source: SSA")+
  ggthemes::theme_pander(base_family = "Fira Sans Condensed", base_size=16)+
  coord_flip()+ # flip axes to make horizontal!
  scale_fill_viridis_d(option = "default")+ # use viridis discrete color palette
  theme(legend.position = "") # hide legend
```

```{r 2-girls}
ggplot(data = top_5_girls_2017)+
  aes(x = reorder(name, n), #note this reorders the x variable from small to large n
      y = percent, # you can use prop if you didn't make a percent variable
      fill = name)+ # optional color!
  geom_col()+
  # all of the above is sufficient, now I'm just making it pretty
  scale_y_continuous(labels = function(x){paste0(x, "%")}, # add percent signs
                     breaks = seq(from = 0, # make line breaks every 0.25%
                                  to = 1.25,
                                  by = 0.25),
                     limits = c(0,1.3), # limit axis to between 0 and 1.2
                     expand = c(0,0))+ # don't let it go beyond this
  labs(x = "Name",
       y = "Percent of All Girls With Name",
       title = "Most Popular Girls Names Since 1880",
       fill = "Girl's Name",
       caption = "Source: SSA")+
  ggthemes::theme_pander(base_family = "Fira Sans Condensed", base_size=16)+
  coord_flip()+ # flip axes to make horizontal!
  scale_fill_viridis_d(option = "default")+ # use viridis discrete color palette
  theme(legend.position = "") # hide legend
```

If you had gone the alternate route by saving an object of names (like I did above with `top_5_boys_2017_alt` and `top_5_girls_2017_alt`), you could filter the data using the `%in%` operator to use for your `data` layer of each plot, like so:

```{r, eval=F}
boys_data <- babynames %>%
  filter(name %in% top_5_boys_2017_alt) # this will only use data for the 5 names

ggplot(data = boys_data) #+... the rest of the plot code above
```

## Question 3 
**Find your name. [If your name isn't in there :(, pick a random name.] `count` by `sex` how many babies since 1880 were named your name. [Hint: if you do this, you'll get the number of *rows* (years) there are in the data. You want to add the number of babies in each row (`n`), so inside `count`, add `wt = n` to weight the count by `n`.] Also add a variable for the percent of each sex.**


<!--WRITE YOUR ANSWERS BELOW -->

```{r 3}
babynames %>%
  filter(name == "Ryan") %>%
  count(sex, wt=n) %>%
  mutate(percent = round((n/sum(n)*100),2))
```

## Question 4
**Make a line graph of the number of babies with your name over time, `color`ed by `sex`.**

<!--WRITE YOUR ANSWERS BELOW -->

```{r 4}
# note here I'm going to wrangle the data and then pipe it directly into ggplot
# you can wrangle the data and save it as a different tibble, then use THAT tibble
# for your (data = ...) command in ggplot

# first wrangle data
babynames %>%
  filter(name == "Ryan") %>%

  # now we pipe into ggplot
  ggplot(data = .)+ # the "." is a placeholder for the stuff above!
  aes(x = year,
      y = n,
      color = sex)+
  geom_line(size = 1)+
  scale_color_manual(values = c("F" = "#e64173", # make my own colors
                                "M" = "#0047AB"))+
  labs(x = "Year",
       y = "Number of Babies",
       title = "Popularity of Babies Named 'Ryan'",
       color = "Sex",
       caption = "Source: SSA")+
    theme_classic(base_family = "Fira Sans Condensed", base_size=16)
```


## Question 5
### Part A
**Find the most common name for boys by year between 1980-2017. [Hint: you'll want to first `group_by(year)`. Once you've got all the right conditions, you'll get a table with a lot of data. You only want to `slice(1)` to keep just the 1st row of each year's data.]**

<!--WRITE YOUR ANSWERS BELOW -->

```{r 5-a}
babynames %>%
  group_by(year) %>% # we want one observation per year
  filter(sex == "M",
         year>1979) %>% # or >==1980
  arrange(desc(n))%>% # start with largest n first
  slice(1) # take first row only
```

### Part B
**Now do the same for girls.**

<!--WRITE YOUR ANSWERS BELOW -->

```{r 5-b}
babynames %>%
  group_by(year) %>% # we want one observation per year
  filter(sex == "F",
         year>1979) %>% # or >==1980
  arrange(desc(n))%>% # start with largest n first
  slice(1) # take first row only
```

## Question 6
**Now let's graph the evolution of the most common names since 1880.**

### Part A
**First, find out what are the top 5 *overall* most popular names for boys and for girls in the data. [Hint: first `group_by(name)`.] You may want to create two objects, each with these top 5 names as character elements.**

<!--WRITE YOUR ANSWERS BELOW -->

```{r 6-a-boys}
babynames %>%
  group_by(name) %>% # we want one row per name
  filter(sex=="M") %>%
  summarize(total=sum(n)) %>% # add upp all of the n's for all years for each name
  arrange(desc(total)) %>% # list largest total first
  slice(1:5) 

# make a vector of the names (we'll need this for our graph below)
top_boys_names<-c("James", "John", "Robert", "Michael", "William")

# you could alternatively add a command, 
# %>% pull(name) to the first chunk of code, 
# and it would do the same thing, but we'd want to save it, 
# for example:

babynames %>%
  group_by(name) %>% # we want one row per name
  filter(sex=="M") %>%
  summarize(total=sum(n)) %>% # add upp all of the n's for all years for each name
  arrange(desc(total)) %>% # list largest total first
  slice(1:5) %>%
  pull(name)
```

```{r 6-a-girls}
babynames %>%
  group_by(name) %>% # we want one row per name
  filter(sex=="F") %>%
  summarize(total=sum(n)) %>% # add upp all of the n's for all years for each name
  arrange(desc(total)) %>% # list largest total first
  slice(1:5)

# make a vector of the names (we'll need this for our graph below)
top_girls_names<-c("Mary", "Elizabeth", "Patricia", "Jennifer", "Linda")
```

### Part B
**Now make two `line`graphs of these 5 names over time, one for boys, and one for girls. [Hint: you'll first want to subset the data to use for your `data` layer in the plot. First `group_by(year)` and also make sure you only use the names you found in Part A. Try using the `%in%` command to do this.]**

<!--WRITE YOUR ANSWERS BELOW -->

```{r 6-b-boys}
babynames %>%
  group_by(year) %>%
  filter(sex == "M",
         name %in% top_boys_names) %>%
  ggplot(data = .,
         aes(x = year,
             y = prop,
             color = name))+
  geom_line(size = 1)+
      labs(x = "Year",
         y = "Proportion of Babies with Name",
         title = "Most Popular Boys Names Since 1880",
         color = "Boy's Name",
         caption = "Source: SSA")+
    theme_classic(base_family = "Fira Sans Condensed", base_size=16)
```

```{r 6-b-girls}
babynames %>%
  group_by(year) %>%
  filter(sex == "F",
         name %in% top_girls_names) %>%
  ggplot(data = .,
         aes(x = year,
             y = prop,
             color = name))+
  geom_line(size = 1)+
    labs(x = "Year",
         y = "Proportion of Babies with Name",
         title = "Most Popular Girls Names Since 1880",
         color = "Girl's Name",
         caption = "Source: SSA")+
    theme_classic(base_family = "Fira Sans Condensed", base_size=16)
```

## Question 7
**Bonus (hard!): What are the 10 most common "gender-neutral" names? [This is hard to define. For our purposes, let's define this as names where between 48% and 52% of the babies with the name are Male.]**

<!--WRITE YOUR ANSWERS BELOW -->

There's a lot to this, so I'll break this up step by step and show you what happens at each major step. 

We want to find the names where 48% to 52% of the babies with the name are male, as I defined in the footnote. First let's `mutate` a variable to figure out how many babies with a particular name are male. 

To do this, we'll need to make a two variables to count the number of `male`s and `female`s of each name each year. We'll use the `ifelse()` function for each:

1. Make a `male` variable where, for each name in each year, if `sex=="M"`, then count the number of males (`n`) that year, otherwise set it equal to `0`.
2. Make a `female` variable where, for each name in each year, if `sex=="F"`, then count the number of females (`n`) that year, otherwise set it equal to `0`.

```{r}
babynames %>%
  mutate(male = ifelse(sex == "M", n, 0),
         female = ifelse(sex == "F", n, 0))
```

Now with this variable, we want to count the total number of males and females with each name over the entire dataset. Let's first `group_by(name)` so we'll get one row for every name. We will `summarize()` and take the `sum` of our `male` and of our `female` variables. 

```{r}
babynames %>%
  mutate(male = ifelse(sex == "M", n, 0),
         female = ifelse(sex == "F", n, 0)) %>%
  group_by(name) %>%
    summarize(Male = sum(male),
              Female = sum(female))
```

Now, we want to figure out what *fraction* of each name is Male or Female. It doesn't matter which we do here, I'll do Male. `mutate()` a new variable I'll call `perc_male` for the percent of the name being for Male babies. It takes the summed variables we made before, and takes the fraction that are Male, multiplying by 100 to get percents (which isn't necessary, but is easy to read).

```{r}
babynames %>%
  mutate(male = ifelse(sex == "M", n, 0),
         female = ifelse(sex == "F", n, 0)) %>%
  group_by(name) %>%
    summarize(Male = sum(male),
              Female = sum(female))%>%
  mutate(perc_male = (Male/(Male+Female)*100))
```

Right now, it's still in alphabetical order. We want to arrange it by `perc_male`, and more importantly, we want `perc_male` to be between 48 and 52, so let's `filter` accordingly: 

```{r}
babynames %>%
  mutate(male = ifelse(sex == "M", n, 0),
         female = ifelse(sex == "F", n, 0)) %>%
  group_by(name) %>%
    summarize(Male = sum(male),
              Female = sum(female))%>%
  mutate(perc_male = (Male/(Male+Female)*100)) %>%
  arrange(perc_male) %>%
  filter(perc_male > 48,
         perc_male < 52)
```

This gives us a lot of names, all falling between 48% and 52% male. But we want the most popular names that are in this range. So let's finally `mutate` a new variable called `total` that simply adds the number of `Male` and `Female` babies with a name. Then let's `arrange` our results by `desc(total)` to get the largest first, and then `slice(1:10)` to get the top 10 only. 

```{r}
babynames %>%
  mutate(male = ifelse(sex == "M", n, 0),
         female = ifelse(sex == "F", n, 0)) %>%
  group_by(name) %>%
    summarize(Male = sum(male),
              Female = sum(female))%>%
  mutate(perc_male = (Male/(Male+Female)*100)) %>%
  arrange(perc_male) %>%
  filter(perc_male > 48,
         perc_male < 52) %>%
  mutate(total = Male+Female) %>%
  arrange(desc(total)) %>%
  slice(1:10)
```

---

# Political and Economic Freedom Around the World

**For the remaining questions, we'll look at the relationship between Economic Freedom and Political Freedom in countries around the world today. Our data for economic freedom comes from the [Fraser Institute](https://www.fraserinstitute.org/economic-freedom/dataset?geozone=world&year=2016&page=dataset), and our data for political freedom comes from [Freedom House](https://freedomhouse.org/content/freedom-world-data-and-resources).**

## Question 8
Download these two datasets that I've cleaned up a bit: [If you want a challenge, try downloading them from the websites and cleaning them up yourself!]

- [<i class="fas fa-table"></i> `econ_freedom.csv`](http://metricsf21.classes.ryansafner.com/data/econ_freedom.csv)
- [<i class="fas fa-table"></i> `pol_freedom.csv`](http://metricsf21.classes.ryansafner.com/data/pol_freedom.csv)

Below is a brief description of the variables I've put in each dataset:

### Econ Freedom

| Variable | Description |
|----------|-------------|
| `year` | Year |
| `ISO` | Three-letter country code |
| `country` | Name of the country |
| `ef_index` | Total economic freedom index (0 - least to 100 - most) |
| `rank` | Rank of the country in terms of economic freedom |
| `continent` | Continent the country is in |

### Pol Freedom

| Variable | Description |
|----------|-------------|
| `country` | Name of the country |
| `C/T` | Whether the location is a country (C) or territory (T) |
| `year` | Year |
| `status` | Whether the location is Free (F), Partly Free (F) or Not Free (NF) |
| `fh_score` | Total political freedom index (0 - least to 100 - most) |

Import and save them each as an object using `my_df_name <- read_csv("name_of_the_file.csv")`. I suggest one as `econ` and the other as `pol`, but it's up to you. Look at each object you've created.

<!--WRITE YOUR ANSWERS BELOW -->

```{r 8}
# import data with read_csv from readr

# note these file paths assume you have a folder called "data" in your working directory

# if you used an R project and did just that (or downloaded my R Project from the website)
# then you already have this done

econ<-read_csv("data/econ_freedom.csv")
pol<-read_csv("data/pol_freedom.csv")

# look at each dataframe
econ
pol
```

## Question 9
**Now let's join them together so that we can have a single dataset to work with. You can learn more about this [in the 1.4 slides](https://metricsf21.classes.ryansafner.com/slides/1.4-slides#143). Since both datasets have both `country` and `year` (spelled exactly the same in both!), we can use these two variables as a `key` to combine observations. Run the following code (substituting whatever you want to name your objects):**

```{r, echo = T, eval = F}
freedom <- left_join(econ, pol, by=c("country", "year")
```

**Take a look at `freedom` to make sure it appears to have worked.**

<!--WRITE YOUR ANSWERS BELOW -->

```{r 9}
freedom <- left_join(econ, pol, by=c("country", "year"))
freedom
```

## Question 11
### Part A
**Make a barplot of the 10 countries with the highest Economic Freedom index score in 2018. You may want to find this first and save it as an object for your plot's `data` layer. Use `geom_col()` since we will map `ef_index` to `y`. If you want to order the bars, set `x = fct_reorder(ISO, desc(ef_index))` to reorder `ISO` (or `country`, if you prefer) by EF score in descending order.**

<!--WRITE YOUR ANSWERS BELOW -->

```{r 10-a}
# grab the top 10 countries by ef in 2018
ef_10<-freedom %>%
  filter(year == 2018) %>%
  arrange(desc(ef_index)) %>%
  slice(1:10)

# look at it just to check
ef_10

# now plot it
ggplot(data = ef_10)+
  aes(x = fct_reorder(ISO, desc(ef_index)), # reorder ISO by ef in order
      y = ef_index)+
  geom_col(aes(fill = continent))+ # coloring is optional
  # above is sufficient, now let's just make it prettier
  geom_text(aes(label = ef_index), # add the score onto the bar
            vjust = 1.2, # adjust it vertically
            color = "white"
            )+
  scale_y_continuous(breaks = seq(0,10,2),
                     limits = c(0,10),
                     expand = c(0,0)
                     )+
  labs(x = "Country",
       y = "Economic Freedom Score",
       title = "Top 10 Countries by Economic Freedom",
       caption = "Source: Frasier Institute",
       fill = "Continent")+
  theme_minimal(base_family = "Fira Sans Condensed")+
  theme(legend.position = "bottom",
        plot.title = element_text(face = "bold", size = rel(1.5))
        )
```

### Part B
**Make a barplot of the 10 countries with the highest Freedom House index score in 2018, similar to what you did for Part A.**

<!--WRITE YOUR ANSWERS BELOW -->

```{r 10-b}
# grab the top 10 countries by fh in 2018
pf_10<-freedom %>%
  filter(year == 2018) %>%
  arrange(desc(fh_score)) %>%
  slice(1:10)

# look at it just to check
pf_10

# now plot it
ggplot(data = pf_10)+
  aes(x = fct_reorder(ISO, desc(fh_score)),
      y = fh_score)+
  geom_col(aes(fill = continent))+ # coloring is optional
  # above is sufficient, now let's just make it prettier
  geom_text(aes(label = fh_score), # add the score onto the bar
            vjust = 1.2, # adjust it vertically
            color = "white")+
  scale_y_continuous(breaks = seq(0,100,20),
                     limits = c(0,100),
                     expand = c(0,0))+
  labs(x = "Country",
       y = "Political Freedom Score",
       title = "Top 10 Countries by Political Freedom",
       caption = "Source: Freedom House",
       fill = "Continent")+
  theme_minimal(base_family = "Fira Sans Condensed")+
  theme(legend.position = "bottom",
        plot.title = element_text(face = "bold", size = rel(1.5))
        )
```

## Question 11
**Now make a scatterplot of Political freedom (`fh_score` as `y`) on Economic Freedom (`ef_index` as `x`) and `color` by `continent`.**

<!--WRITE YOUR ANSWERS BELOW -->

```{r 11}
ggplot(data = freedom)+
  aes(x = ef_index,
      y = fh_score)+
  # doing just geom_point() is fine, but since there's a lot of overlap, here are some things I like to do:
  geom_point(aes(fill = continent), # fill the points with color by continent
             alpha = 0.9, # make points slightly transparent
             color = "white", # outline the points with a white border
             pch = 21, # this shape has an outline and a fill color
             size = 3)+
  scale_x_continuous(breaks = seq(0,10,2),
                     limits = c(0,10),
                     expand = c(0,0))+
  scale_y_continuous(breaks = seq(0,100,20),
                     limits = c(0,105),
                     expand = c(0,0))+
  labs(x = "Economic Freedom Score",
       y = "Political Freedom Score",
       caption = "Sources: Frasier Institute, Freedom House",
       title = "Economic Freedom & Political Freedom",
       fill = "Continent")+
  theme_minimal(base_family = "Fira Sans Condensed")+
  theme(legend.position = "bottom",
        plot.title = element_text(face = "bold", size = rel(1.5))
        )
```

Note, I meant to ask you to look at one year only, e.g. 2018. We would just have to filter first:

```{r 11-2018}
# save as p
p <- freedom %>%
  filter(year == "2018") %>%
ggplot(data = .)+
  aes(x = ef_index,
      y = fh_score)+
  geom_point(aes(fill = continent),
             alpha = 0.9, 
             color = "white",
             pch = 21, 
             size = 3)+
  scale_x_continuous(breaks = seq(0,10,2),
                     limits = c(0,10),
                     expand = c(0,0))+
  scale_y_continuous(breaks = seq(0,100,20),
                     limits = c(0,105),
                     expand = c(0,0))+
  labs(x = "Economic Freedom Score",
       y = "Political Freedom Score",
       caption = "Sources: Frasier Institute, Freedom House",
       title = "Economic Freedom & Political Freedom",
       fill = "Continent")+
  theme_minimal(base_family = "Fira Sans Condensed")+
  theme(legend.position = "bottom",
        plot.title = element_text(face = "bold", size = rel(1.5))
        )

# look at it
p
```


## Question 12
**Save your plot from Question 11 as an object, and add a new layer where we will highlight a few countries. Pick a few countries (I suggest using the `ISO` code) and create a new object `filtering` the data to only include these countries (again the `%in%` command will be most helpful here).**

**Additionally, _install_ and _load_ a package called `"ggrepel"`, which will adjust labels so they do not overlap on a plot.**

**Then, add the following layer to your plot:**

```{r, eval = F}
geom_label_repel(data = countries, # or whatever object name you created
                     aes(x = ef_index,
                         y = fh_score,
                         label = ISO, # show ISO as label (you could do country instead)
                         color = continent),
                     alpha = 0.5, # make it a bit transparent
                     box.padding = 0.75, # control how far labels are from points
                     show.legend = F) # don't want this to add to the legend
```

**This should highlight these countries on your plot.**

<!--WRITE YOUR ANSWERS BELOW -->

```{r 12}
library(ggrepel)

# pick some countries
some_countries <- freedom %>%
  filter(year==2018,
         country %in% c("United States",
                        "United Kingdom",
                        "Sweden",
                        "China",
                        "Singapore",
                        "Russian Federation",
                        "Korea, Rep.",
                        "Hong Kong SAR, China"))
# add layer 
p + geom_label_repel(data = some_countries, # or whatever object name you created
                     aes(x = ef_index,
                         y = fh_score,
                         label = ISO, # show ISO as label (you could do country instead)
                         color = continent),
                     alpha = 0.75, # make it a bit transparent
                     box.padding = 0.75, # control how far labels are from points
                     show.legend = F) # don't want this to add to the legend
```

## Question 13
**Let's just look only at the United States and see how it has fared in both measures of freedom over time. `filter()` the data to look only at `ISO == "USA"`. Use both a `geom_point()` layer and a `geom_path()` layer, which will connect the dots over time. Let's also see this by labeling the years with an additional layer `geom_text_repel(aes(label = year))`.**

<!--WRITE YOUR ANSWERS BELOW -->

```{r 13}
# save plot as us
us<-freedom %>%
  filter(ISO == "USA") %>%
ggplot(data = .)+
  aes(x = ef_index,
      y = fh_score)+
  geom_point(color = "red")+
  geom_path(color = "red")+
  geom_text_repel(aes(label = year),
                  color = "red")+
  scale_x_continuous(breaks = seq(8,8.5,0.05),
                     limits = c(8,8.5),
                     expand = c(0,0))+
  scale_y_continuous(breaks = seq(85,95,1),
                     limits = c(85,95),
                     expand = c(0,0))+
  labs(x = "Economic Freedom Score",
       y = "Political Freedom Score",
       caption = "Sources: Frasier Institute, Freedom House",
       title = "U.S. Political & Economic Freedom, 2013—2018",
       fill = "Continent")+
  theme_minimal(base_family = "Fira Sans Condensed")+
  theme(legend.position = "bottom",
        plot.title = element_text(face = "bold", size = rel(1.5))
        )
# look at it
us
```

Note that the way I zoomed in on the scales, these look like pretty dramatic changes!

If we maintain the full perspective, the change appears minor. Be very careful how you present your analysis!

```{r 13-zoomed-out}
us+
  scale_x_continuous(breaks = seq(0,10,1),
                     limits = c(0,10),
                     expand = c(0,0)
                     )+
  scale_y_continuous(breaks = seq(0,100,10),
                     limits = c(0,100),
                     expand = c(0,0)
                     )
```
