---
title: "Problem Set 3"
author: "YOUR NAME HERE" # put your name here! 
date: "ECON 480 — Fall 2021"
output: html_document # change to pdf_document if you'd like, this will make a webpage (you can email it and open in any browser)
---

<!--CLICK "KNIT" ABOVE TO RENDER TO HTML, PDF, OR WORD OUTPUT
In fact, try knitting right away and see what this produces!

To knit to PDF, you will need to install a LaTeX distribution, which you can do
easily by running the following code in R:

install.packages("tinytex")
tinytex::install_tinytex()
-->

*Due by Thursday October 7*

# Instructions

For this problem set, you may submit handwritten answers on a plain sheet of paper, or download and type/handwrite on the PDF.

Alternatively, you may download the `.Rmd` file, do the homework in markdown, and email to me a single `knit`ted `html` or `pdf` file (and be sure that it shows all of your code). 

You may work together (and I highly encourage that) but you must turn in your own answers. I grade homeworks 70% for completion, and for the remaining 30%, pick one question to grade for accuracy - so it is best that you try every problem, even if you are unsure how to complete it accurately.

# Theory and Concepts

## Question 1
**In your own words, describe what exogeneity and endogeneity mean, and how they are related to bias in our regression. What things can we learn about the bias if we know $X$ is endogenous?**

<!--WRITE YOUR ANSWERS BELOW -->


## Question 2
**In your own words, describe what $R^2$ means. How do we calculate it, what does it tell us, and how do we interpret it?**

<!--WRITE YOUR ANSWERS BELOW -->


## Question 3

**In your own words, describe what the standard error of the regression ($SER$) means. How do we calculate it, what does it tell us, and how do we interpret it?**

<!--WRITE YOUR ANSWERS BELOW -->

## Question 4

**In your own words, describe what homoskedasticity and heteroskedasticity mean: both in ordinary English, and in terms of the graph of the OLS regression line.**

<!--WRITE YOUR ANSWERS BELOW -->

## Question 5

**In your own words, describe what the variation in $\hat{\beta_1}$ (either variance or standard error) means, or is measuring. What three things determine the variation, and in what way?**

<!--WRITE YOUR ANSWERS BELOW -->

## Question 6

**In your own words, describe what a $p$-value means, and how it is used to establish statistical significance.**

<!--WRITE YOUR ANSWERS BELOW -->

## Question 7

**A researcher is interested in examining the impact of illegal music downloads on commercial music sales. The author collects data on commercial sales of the top 500 singles from 2017 ($Y$) and the number of downloads from a web site that allows `file sharing' ($X$). The author estimates the following model**

$$\text{music sales}_i = \beta_0+\beta_1 \text{illegal downloads}_i + u_i$$

**The author finds a large, positive, and statistically significant estimate of $\hat{\beta_1}$. The author concludes these results demonstrate that illegal downloads actually *boost* music sales. Is this an unbiased estimate of the impact of illegal music on sales? Why or why not? Do you expect the estimate to overstate or understate the true relationship between illegal downloads and sales?**

<!--WRITE YOUR ANSWERS BELOW -->

## Question 8

**A pharmaceutical company is interested in estimating the impact of a new drug on cholesterol levels. They enroll 200 people in a clinical trial. People are randomly assigned the treatment group or into the control group. Half of the people are given the new drug and half the people are given a sugar pill with no active ingredient. To examine the impact of dosage on reductions in cholesterol levels, the authors of the study regress the following model:**

$$\text{cholesterol level}_i = \beta_0+\beta_1 \text{dosage level}_i + u_i$$

**For people in the control group, dosage level$_i=0$ and for people in the treatment group, dosage level$_i$ measures milligrams of the active ingredient. In this case, the authors find a large, negative, statistically significant estimate of $\hat{\beta_1}$. Is this an unbiased estimate of the impact of dosage on change in cholesterol level? Why or why not? Do you expect the estimate to overstate or understate the true relationship between dosage and cholesterol level?**

<!--WRITE YOUR ANSWERS BELOW -->

# Theory Problems

**For the following questions, please *show all work* and explain answers as necessary. You may lose points if you only write the correct answer. You may use `R` to *verify* your answers, but you are expected to reach the answers in this section "manually."**

## Question 9

**A researcher wants to estimate the relationship between average weekly earnings $(AWE$, measured in dollars) and $Age$ (measured in years) using a simple OLS model. Using a random sample of college-educated full-time workers aged 25-65 yields the following:**

$$\widehat{AWE} = 696.70+9.60 \, Age$$

### Part A

**Interpret what $\hat{\beta_0}$ means in this context.**

<!--WRITE YOUR ANSWERS BELOW -->

### Part B

**Interpret what $\hat{\beta_1}$ means in this context.**

<!--WRITE YOUR ANSWERS BELOW -->

### Part C

**The $R^2=0.023$ for this regression. What are the units of the $R^2$, and what does this mean?**

<!--WRITE YOUR ANSWERS BELOW -->

### Part D

**The $SER, \, \hat{\sigma_u}=624.1$ for this regression. What are the units of the SER in this context, and what does it mean? Is the SER large in the context of this regression?**

<!--WRITE YOUR ANSWERS BELOW -->

### Part E

**Suppose Maria is 20 years old. What is her predicted $\widehat{AWE}$?**

<!--WRITE YOUR ANSWERS BELOW -->

### Part F

**Suppose the data shows her *actual* $AWE$ is $430. What is her residual? Is this a relatively good or a bad prediction?^[Hint: compare your answer here to your answer in Part D.]**

<!--WRITE YOUR ANSWERS BELOW -->

### Part G

**What does the error term, $\hat{u_i}$ represent in this case? What might individuals have different values of $u_i$?**

<!--WRITE YOUR ANSWERS BELOW -->

### Part H

**Do you think that $Age$ is exogenous? Why or why not? Would we expect $\hat{\beta_1}$ to be too *large* or too *small*?**

<!--WRITE YOUR ANSWERS BELOW -->

## Question 10

**Suppose a researcher is interested in estimating a simple linear regression model:**

$$Y_i=\beta_0+\beta_1X_i+u_i$$
**In a sample of 48 observations, she generates the following descriptive statistics:**

- $\bar{X}=30$
- $\bar{Y}=63$
- $\displaystyle\sum^n_{i=1}(X_i-\bar{X})^2= 6900$
- $\displaystyle\sum^n_{i=1}(Y_i-\bar{Y})^2= 29000$
- $\displaystyle\sum^n_{i=1}(X_i-\bar{X})(Y_i-\bar{Y})=13800$
- $\displaystyle\sum^n_{i=1}\hat{u}^2=1656$

### Part A

**What is the OLS estimate of $\hat{\beta_1}$?**

<!--WRITE YOUR ANSWERS BELOW -->

### Part B

**What is the OLS estimate of $\hat{\beta_0}$?**

<!--WRITE YOUR ANSWERS BELOW -->

### Part C

**Suppose the OLS estimate of $\hat{\beta_1}$ has a standard error of $0.072$. Could we probably reject a null hypothesis of $H_0: \beta_1=0$ at the 95% level?**

<!--WRITE YOUR ANSWERS BELOW -->

### Part D

**Calculate the $R^2$ for this model. How much variation in $Y$ is explained by our model?**

<!--WRITE YOUR ANSWERS BELOW -->

### Part E

**How large is the average residual?**

<!--WRITE YOUR ANSWERS BELOW -->

# R Questions

Answer the following questions using `R`. When necessary, please write answers in the same document (knitted `Rmd` to `html` or `pdf`, typed `.doc(x)`, or handwritten) as your answers to the above questions. Be sure to include (email or print an `.R` file, or show in your knitted `markdown`) your code and the outputs of your code with the rest of your answers.

## Question 11

- [<i class="fas fa-table"></i> `mlbattend.csv`](http://metricsf20.classes.ryansafner.com/data/mlbattend.csv)

Download the `MLBattend` dataset. This data contains data on attendance at major league baseball games for all 32 MLB teams from the 1970s-2000. We want to answer the following question:

> "How big is home-field advantage in baseball? Does a team with higher attendance at home games over their season have score more runs over their season?"

### Part A

**Clean up the data a bit by `mutate()`-ing a variable to measure home attendance in millions. This will make it easier to interpret your regression later on.**

<!--WRITE YOUR ANSWERS BELOW -->

```{r}
# write your code here
```

### Part B

**Get the correlation between Runs Scored and Home Attendance.**

<!--WRITE YOUR ANSWERS BELOW -->

```{r}
# write your code here
```


**Plot a scatterplot of Runs Scored (`y`) on Home Attendance (`x`). Add a regression line.**

<!--WRITE YOUR ANSWERS BELOW -->

```{r}
# write your code here
```

### Part D

**We want to estimate a regression of Runs Scored on Home Attendance:**

$$
\widehat{\text{runs scored}_i} = \beta_0 + \beta_1 \text{home attendance}_i
$$

**Run this regression in `R`.**

**What are $\hat{\beta_0}$ and $\hat{\beta_1}$ for this model? Interpret them in the context of our question. [Hint: make sure to save your regression model as an object, and get a `summary()` of it. This object will be needed later.]**

<!--WRITE YOUR ANSWERS BELOW -->

```{r}
# write your code here
```

### Part E

**Write out the estimated regression equation.**

<!--WRITE YOUR ANSWERS BELOW -->

```{r}
# write your code here
```

### Part F

**Make a regression table of the output (using the `huxtable` package).**

<!--WRITE YOUR ANSWERS BELOW -->

```{r}
# write your code here
```

### Part G

**Check the goodness of fit statistics. What is the $R^2$ and the SER of this model? Interpret them both in the context of our question.**

<!--WRITE YOUR ANSWERS BELOW -->

```{r}
# write your code here
```

### Part H

**Now let's start running some diagnostics of the regression. Make a histogram of the residuals. Do they look roughly normal? [Hint: you will need to use the `broom` package's `augment()` command on your saved regression object to add containing the residuals (`.resid`), and save this as a new object - to be your data source for the plot in this question and the next question.]**

<!--WRITE YOUR ANSWERS BELOW -->

```{r}
# write your code here
```

### Part I

**Make a residual plot.**

<!--WRITE YOUR ANSWERS BELOW -->

```{r}
# write your code here
```

### Part J

**Test the regression for heteroskedasticity. Are the errors homoskedastic or heteroskedastic?**

**[Hint: use the `lmtest` package's `bptest()` command on your saved regression object.]**

**Run another regression using robust standard errors. [Hint: use the `estimatr` package's `lm_robust()` command and save the output like the following:**

```{r, eval = F}
reg_robust <-lm_robust(y ~ x, data = the_data, # change y, x, and data names to yours
                              se_type = "stata") # we'll use this method to calculate
```

**Now make another regression output table with `huxtable`, with one column using regular standard errors (just use your original saved regression object) and another using robust standard errors (use this new saved object).**

<!--WRITE YOUR ANSWERS BELOW -->

```{r}
# write your code here
```

### Part K

**Test the data for outliers. If there are any, identify which team(s) and season(s) are outliers. [Hint: use the `car` package's `outlierTest()` command on your saved regression object.]**

<!--WRITE YOUR ANSWERS BELOW -->

```{r}
# write your code here
```

### Part L

**Look back at your regression results. What is the marginal effect of home attendance on runs scored? Is this statistically significant? Why or why not?**

<!--WRITE YOUR ANSWERS BELOW -->

```{r}
# write your code here
```


### Part M

**Now we'll try out the `infer` package to understand the $p$-value for our observed slope in our regression model.**

**First, save the (value of) our sample $\hat{\beta_1}$ from your regression in Part D as an object, I suggest:**

```{r, eval = F}
our_slope = 123 # replace "123" with whatever number you found for the slope in part D
```

**Then, install and load the `infer` package, and then run the following simulation:**

```{r, eval = F}
# save our simulations as an object (I called it "sims")
sims <- data %>% # "data" here is whatever you named your dataframe!
  specify(y ~ x) %>% # replacing y and x with your variable names
  hypothesize(null = "independence") %>% # H_0 is that slope is 0, x and y are independent
  generate(reps = 1000,
           type = "permute") %>% # make 1000 samples assuming H_0 is true
  calculate(stat = "slope") # estimate slope in each sample

# look at it
sims

# calculate p value
sims %>%
  get_p_value(obs_stat = our_slope,
              direction = "both") # a two-sided H_a: slope =/= 0
```

**Compare to the $p$-value in your original regression output in previous parts of this question.**

<!--WRITE YOUR ANSWERS BELOW -->

```{r}
# write your code here
```

### Part N

**Make a histogram of the simulated slopes, and plot our sample slope on that histogram, shading the $p$-value.** 

**[You can pipe `sims` into `visualize(obs_stat = our_slope)`, or use `ggplot2` to plot a histogram in the normal way, using `sims` as the data source and add a `geom_vline(xintercept = our_slope)` to show our finding on the distribution.]**

<!--WRITE YOUR ANSWERS BELOW -->

```{r}
# write your code here
```
