---
title: "2.7 — Inference for Regression - R Practice"
draft: false
date: "2020-06-08"
linktitle: "2.7 — Inference for Regression"
menu:
  r:
    parent: R Practice
    weight: 6
type: docs
output:
  blogdown::html_page:
    toc: false
---

```{r message=FALSE, warning=FALSE, include=FALSE}
library(tidyverse)
library(pander)
```

# Set Up
To minimize confusion, I suggest creating a new `R Project` (e.g. `infer_practice`) and storing any data in that folder on your computer.

I have already made an R project you can download (as a `.zip`), unzip, and open the `.Rproj` file in R Studio, or there is an R project you can use on the cloud:

<p style="text-align:center"><a target="_blank" href="/r/2.7-r.zip"><span class="btn btn-primary btn-lg"><i class="fas fa-file-archive"></i> R Project</a></span> <a target="_blank" href="https://rstudio.cloud/spaces/163934/project/2990242"><span class="btn btn-primary btn-lg"><i class="fas fa-cloud"></i> R Studio Cloud</a></span></p>

## Question 1

Let’s use the `diamonds` data built into `ggplot`. Simply load `tidyverse` and then to be explicit, save this as a tibble (feel free to rename it) with `diamonds <- diamonds`.

Let's answer the following questions:

> What is the effect of carat size on a diamond's price?

### Part A

Just to see what we're looking at, make a scatterplot using `ggplot()`, with `x` as `carat` and `y` as `price`, and add a regression line.

## Question 2

Suppose we want to estimate the following relationship:

$$\widehat{\text{price}}_i = \beta_0 + \beta_1 \text{carat}_i + u_i$$

Run a regression of `price` on `carat` using `lm()` and get a `summary()` of it. Be sure to save your regression model as an object, we'll need it later.

### Part A

Write out the estimated regression equation.

### Part B

Make a regression table of the output (using the `huxtable` package).

### Part C

What is $\hat{\beta_1}$ for this model? Interpret it in the context of our question.

### Part D

Use the `broom` package's `tidy()` command on your regression object, and calculate confidence intervals for your estimates by setting `conf.int = T` inside `tidy()`. 

What is the 95% confidence interval for $\hat{\beta_1}$, and what does it mean?

Save these endpoints as an object (either by taking your `tidy()`-ed regression and `filter()`-ing the `term == carat` and then `select()`-ing the columns with the confidence interval and then saving this; or simply assigning the two values as a vector, `c( , )`, to an object).

### Part E

Save your estimated $\hat{\beta_1}$ as an object, we'll need it later with `infer` (either by taking your `tidy()`-ed regression and `filter()`-ing the `term == carat` and then `select()`-ing the column with the `estimate` and then saving this; or simply assigning the values to an object).

## Question 3

Now let’s use `infer`. Install it if you don’t have it, then load it.

### Part A

Let’s generate a confidence interval for $\hat{\beta_1}$ by simulating the sampling distribution of $\hat{\beta_1}$. Run the following code, which will `specify()` the model relationship and `generate()` 1,000 repetitions using the `bootstrap` method of resampling data points randomly from our sample, with replacement.

```{r, eval = F}
# save our simulations as an object (I called it "boot")
boot <- diamonds %>% # or whatever you named your dataframe!
  specify(carat ~ price) %>% # our regression model
  generate(reps = 1000, # run 1000 simulations
           type = "bootstrap") %>% # using bootstrap method
  calculate(stat = "slope") # estimate slope in each simulation

# look at it
boot
```

Note this will take a few minutes, its doing a lot of calculations! What does it show you when you look at it?

### Part B

Continue by piping your object from Part A into `get_confidence_interval()`. Set `level = 0.95, type = "se"` and `point_estimate` equal to our estimated $\hat{\beta_1}$ (saved) from Question 2 Part E.

```{r, eval = F}
boot +
  get_confidence_interval(level = 0.95,
                          type = "se", 
                          point_estimate = beta_1_hat) # or whatever you saved it as
```

### Part C

Now instead of `get_confidence_interval()`, pipe your object from Part A into `visualize()` to see the sampling distribution of $\hat{\beta_1}$ we simulated. You can add `+ shade_ci(endpoints = ...)` setting the argument equal to whatever you called your object containing the confidence interval from Question 2 Part D (I have it named here as `CI_endpoints`).

```{r, eval = F}
boot +
  visualize()+
  shade_ci(endpoints = CI_endpoints) # or whatever you saved it as
```

Compare your *simulated* confidence interval with the *theoretically-constructed* confidence interval from the output of `summary`, and/or of `tidy()` from Question 2.

## Question 4

Now let’s test the following hypothesis:

$$\begin{align*}
H_0: \beta_1 &= 0\\
H_1: \beta_1 &\neq 0\\
\end{align*}$$

### Part A

What does the output of `summary`, and/or of `tidy()` from Question 2 tell you?

### Part B

Let’s now do run this hypothesis test with `infer`, which will simulate the sampling distribution under the null hypothesis that $\beta_1 = 0$. Run the following code, which will `specify()` the model relationship and `hypothesize()` the null hypothesis that there is no relationship between $X$ and $Y$ (i.e. $\beta_1=0)$, and `generate()` 1,000 repetitions using the `permute` method, which will center the distribution at 0, and then  `calculate(stat = "slope")`.

```{r, eval = F}
# save our simulations as an object (I called it "test_sims")
test_sims <- diamonds %>% # or whatever you named your dataframe!
  specify(carat ~ price) %>% # our regression model
  hypothesize(null = "independence") %>% # H_0 is that slope is 0
  generate(reps = 1000, # run 1000 simulations
           type = "permute") %>% # using permute method, centering distr at 0
  calculate(stat = "slope") # estimate slope in each simulation

# look at it
test_sims
```

Note this may also take a few minutes. What does it show you?

### Part C

Pipe your object from the previous part into the following code, which will `get_p_value()`. Inside this function, we are setting `obs_stat` equal to our $\hat{\beta_1}$ we found (from Question 2 part E), and set `direction = "both"` to run a *two*-sided test, since our alternative hypothesis is two-sided, $H_1: \beta_1 \neq 0$.

```{r, eval = F}
test_sims %>%
  get_p_value(obs_stat = beta_1_hat,
              direction = "both")
```

Note the warning message that comes up!

### Part D

Instead of `get_p_value()`, pipe your object from Part B into the following code, which will `visualize()` the null distribution, and (in the second command), place our finding on it and `shade_p_value()`.

```{r, eval =F}
test_sims %>%
  visualize()
```

```{r, eval = F}
test_sims %>%
  visualize()+
  shade_p_value(obs_stat = beta_1_hat, # or whatever you saved it as
                direction = "both") # for two-sided test
```

### Part E

Compare your *simulated* p-value with the *theoretically-constructed* p-value from the output of `summary`, and/or of `tidy()` from Question 2.

Both `summary` and `tidy()` also report the $t$-statistic (`t value` or `statistic`) on this test for `carat` (by default, that $H_0: \beta_1=0)$. What is the estimated test statistic for this model, and what does this number mean? Try to calculate it yourself with the formula:

$$t = \frac{\text{estimate} - \text{null hypothesis value}}{\text{standard error of estimate}}$$

The p-value is the probability of a $t$ statistic at least as large as ours if the null hypothesis were true. `R` constructs a $t$-distribution with `n-k-1` degrees of freedom (`n` is number of observations, `k` is number of $X$-variables) and calculates the probability in the tails of the distribution beyond this $t$ value. You can calculate it yourself (for a two-sided test) with: 

```{r, eval = F}
2 * pt(your_t, # put your t-statistic here
       df = your_df, # put the df number here
       lower.tail = F) # we'll use the right tail
```