---
title: "3.2 — Causal Inference & DAGs — R Practice"
draft: false
date: "2020-06-08"
linktitle: "3.2 — Causal Inference & DAGs"
menu:
  r:
    parent: R Practice
    weight: 7
type: docs
output:
  blogdown::html_page:
    toc: false
---

## Answers

<p style="text-align:center"><a target="_blank" href="/r/3.2-r-answers.html"><span class="btn btn-primary btn-lg"><i class="fas fa-external-link-square-alt"></i> Answers (html)</a> <a target="_blank" href="/r/3.2-r-answers.Rmd"><span class="btn btn-primary btn-lg"><i class="fab fa-markdown"></i> Answers (Rmd)</a></span></p>

For each of the following examples:

i. Write out *all* of the causal pathways from `X` (treatment of interest) to `Y` (outcome of interest).
ii. Identify which variable(s) need to be controlled to estimate the causal effect of `X` on `Y`. You can use [`dagitty.net`](http://dagitty.net) to help you, but you should start trying to recognize these on your own!
iii. Draw the DAGs in `r` using `ggdag`. After setting up the dag with `dagify()` (and specifying `exposure` and `outcome` inside `dagify`), pipe that into `ggdag()`. Try again piping it instead into `ggdag_status()` (to highlight what is X and what is Y). Try again piping it instead into `ggdag_adjustment_set()` to show what needs to be controlled.

Don’t forget to install `ggdag`!

```{r, echo=F, message=F, warning=F}
library(tidyverse)
library(ggdag) # load to draw DAGs! 
```

## Question 1

```{r, echo=F, fig.retina=3}
dagify(Y ~ X + Z,
       X ~ Z,
       exposure = "X",
       outcome = "Y") %>% 
  ggdag(stylized = FALSE, seed = 1)+
  theme_dag_blank()+
  theme(legend.position = "none")
```

## Question 2

```{r, echo=F, fig.retina=3}
dagify(Y ~ X + M,
       M ~ X,
       exposure = "X",
       outcome = "Y") %>% 
  ggdag(stylized = FALSE, seed = 1)+
  theme_dag_blank()+
  theme(legend.position = "none")
```

## Question 3

```{r, echo=F, fig.retina=3}
dagify(Y ~ X + Z + B,
       Z ~ B + A,
       X ~ A,
       exposure = "X",
       outcome = "Y") %>% 
  ggdag(stylized = FALSE, seed = 1)+
  theme_dag_blank()+
  theme(legend.position = "none")
```

## Question 4

```{r, echo=F, fig.retina=3}
dagify(Y ~ X + B + C,
       Z ~ B + A,
       X ~ A,
       C ~ X,
       exposure = "X",
       outcome = "Y") %>% 
  ggdag(stylized = FALSE, seed = 2)+
  theme_dag_blank()+
  theme(legend.position = "none")
```

## Question 5

```{r, echo=F, fig.retina=3}
dagify(Y ~ X + Z + A,
       Z ~ X + A + B,
       B ~ A,
       exposure = "X",
       outcome = "Y") %>% 
  ggdag(stylized = FALSE, seed=2)+
  theme_dag_blank()+
  theme(legend.position = "none")
```