---
title: "Example Presentation"
author: "Ryan Safner"
date: "ECON 480 — Fall 2021"
output:
  beamer_presentation:
    dev: cairo_pdf # Allows different fonts
    latex_engine: xelatex # needed for different fonts
    theme: "metropolis" 

# LaTeX packages needed for huxtable to render tables:
header_includes: 
  - \usepackage{booktabs}
  - \usepackage{array}
  - \usepackage{caption}
  - \usepackage{graphicx}
  - \usepackage{siunitx}
  - \usepackage[normalem]{ulem}
  - \usepackage{colortbl}
  - \usepackage{multirow}
  - \usepackage{hhline}
  - \usepackage{calc}
  - \usepackage{tabularx}
  - \usepackage{threeparttable}
  - \usepackage{wrapfig}
  - \usepackage{adjustbox}
  - \usepackage{hyperref}

classoption: aspectratio=169
---

<!-- If you're looking at this source code, this is a “Beamer” presentation
that uses LaTeX to make PDF slides. 

-->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      message = F,
                      warning = F,
                      fig.retina = 3)
library(tidyverse)
library(knitr)
```

# Overview

- As a student writing an empirical research paper, does writing a longer paper attain a higher grade?
- Simple OLS regression
- Sample data collected from previous classes with paper assignments 
- Strong positive effect: for every marginal page written, grades improve by about 1-2 points 
- Robust to different models 

# Theory

- Dilemma:
1. Longer papers might imply students work hard and write a good paper 
2. But students might also put in low quality filler hoping to inflate their grades 

# Data

- I collected data from 7 different classes at 2 different colleges where I assigned a longer term paper 

```{r, load-data}
papers<-read_csv("../data/paperlengthsregcsv.csv")
source("../files/summaries.R")
papers<-papers %>%
  mutate(Sex=as.factor(Sex),
         Sex=recode(Sex, `F` = "Female", M = "Male")) %>%
  mutate(Female = ifelse(Sex=="Female",1,0),
         Morning = ifelse(Time=="Morning",1,0),
         Hood=ifelse(School=="Hood",1,0),
         Econometrics=ifelse(Class=="Econometrics",1,0),
         Covid=ifelse(Covid=="Yes",1,0))
```

# Variables

| Variable | Description |
|----------|-------------|
| Grade    | Grade on paper assignment (0-100) |
| Pages    | Number of pages written |
| Final | Final course grade for student |
| Gender | Gender of student |
| Class | Class in which paper was assigned |
| School | School of class taught |
| Year | Year of class |
| Time | Time of day class met |
| Covid | Course during Covid? |

# Summary Statistics


```{r}
summary_table(papers, Pages, Grade, Final, Year, Female, Morning, Hood, Econometrics, Covid) %>%
  knitr::kable(., format="latex")
```

# Data: Histogram of X

```{r, fig.retina=3, fig.align="center", out.width="50%"}
papers<-papers %>%
  mutate(Metrics = ifelse(Econometrics==1, "Econometrics", "Other"))


pages_hist<-ggplot(data = papers)+
  aes(x = Pages)+
  geom_histogram(color="white", fill="#e64173",breaks=seq(0,24,2))+
  scale_x_continuous(breaks=seq(0,24,2),
                     limits = c(0,25),
                     expand = c(0,0))+
  scale_y_continuous(breaks=seq(0,50,10),
                     limits = c(0,52),
                     expand = c(0,0))+
  labs(x = "Number of Pages Written",
       y = "Number of Papers")+
    ggthemes::theme_pander(base_family = "Fira Sans Condensed",
           base_size=16)
pages_hist
```

# Data

```{r, fig.retina=3, fig.align="center", out.width="50%"}

pages_hist+facet_wrap(~Metrics)
```


# Data

```{r, fig.retina=3, fig.align="center", out.width="50%"}
scatter_no0s<-ggplot(data = subset(papers, Grade>0))+
  aes(x = Pages,
      y = Grade)+
  geom_jitter(aes(color = Class))+
  geom_smooth(method="lm", color="black")+
  scale_x_continuous(breaks=seq(0,24,2),
                     limits=c(0,25),
                     expand=c(0,0))+
  scale_y_continuous(breaks=seq(0,100,10),
                     limits=c(0,110),
                     expand=c(0,0))+
  labs(x = "Number of Pages Written",
       y = "Paper Grade",
       title = "Pages Written vs. Paper Grade (No 0’s)")+
    ggthemes::theme_pander(base_family = "Fira Sans Condensed",
           base_size=16)+
  theme(legend.position = "bottom")
scatter_no0s
```

# Data

```{r, fig.retina=3, fig.align="center", out.width="50%"}
ggplot(data = subset(papers, Class=="Econometrics"))+
  aes(x = Pages,
      y = Grade)+
  geom_jitter(color = "blue")+
  geom_smooth(method="lm", color="red")+
  scale_x_continuous(breaks=seq(0,24,2),
                     limits=c(0,25),
                     expand=c(0,0))+
  scale_y_continuous(breaks=seq(0,100,10),
                     limits=c(0,110),
                     expand=c(0,0))+
  labs(x = "Number of Pages Written",
       y = "Paper Grade",
       title = "Pages Written vs. Paper Grade (Econometrics Only)")+
    ggthemes::theme_pander(base_family = "Fira Sans Condensed",
           base_size=16)
```

# Empirical Model

$$\begin{aligned}
\text{Paper Grade}_i=& \, \beta_0+\beta_1\text{Paper Length}_i+\beta_2\text{Course Grade}_i\\
&+\beta_3\text{Gender}_i+\beta_4\text{School}_i+\beta_5\text{Covid}_i\\
&+\beta_6\text{Course}_i+u_i\\ \end{aligned}$$

# Results

```{r}
papers0<-papers %>%
  filter(Grade>0)

hood<-papers %>%
  filter(School=="Hood")

metrics<-papers %>%
  filter(Class=="Econometrics")

basicreg<-lm(Grade~Pages, data=papers)
no0reg<-lm(Grade~Pages, data=papers0)
basicmetricsreg<-lm(Grade~Pages, data=metrics)
controlsreg<-lm(Grade~Pages+Final+Female+Hood+Metrics+Covid, data=papers0)
hoodreg<-lm(Grade~Pages+Final+Female+Metrics+Covid, data=hood)
metricsreg<-lm(Grade~Pages+Final+Female+Covid, data=metrics)
```

```{r, results = "asis"}
# ^ Need results “asis” to print a table in LaTeX 
library(huxtable)
huxreg("Baseline" = basicreg,
       "No Os" = no0reg,
       "Econometrics Only" = basicmetricsreg,
       "With Controls" = controlsreg,
       "Hood Only" = hoodreg,
       "Econometrics Only" = metricsreg,
       coefs = c("Constant" = "(Intercept)",
                 "Length" = "Pages",
                 "Course Grade" = "Final",
                 "Hood College" = "Hood",
                 "Female" = "Female",
                 "Econometrics Course" = "MetricsOther",
                 "During Covid?" = "Covid"),
       statistics = c("N" = "nobs",
                      "R-Squared" = "r.squared",
                      "SER" = "sigma"),
       number_format = 2) %>%
  print_latex
```

# Implications

- For every additional page written, we can expect a paper's grade to increase by about a point or less. 

- For econometrics only, marginal effect is even smaller, only less than half of a point increase for every additional page written. 

- Likely endogeneity of length due to unobserved factors such as topic and quality of writing

- *It would be poor advice to recommend students simply to write long papers to earn a higher grade.*
