---
date: "2019-05-05T00:00:00+01:00"
draft: false
linktitle: Data Wrangling
menu:
  resources:
    parent: R Resources
title: "Data Wrangling Tips & Tricks"
toc: true
type: docs
weight: 3
---

```{r}
library(tidyverse)
```

## Importing Data

The `janitor` package 

`cleannames()`

- generally makes lowercase names

## Dealing with Missing Data

When calculating statistics (e.g. with `summarize()`), many calculations will give errors if your data contains `NA`s. 

### Example: Calculating Mean

```{r, echo = T}
data_missing <- tribble(
  ~x, ~y,
  2, 3,
  1, 4,
  NA, 2,
  3, NA,
  7, 8
)
```

Now if we were to get the mean of x: 
```{r}
data_missing %>% 
  summarize(mean_x = mean(x))
```

It gives us `NA`. 

One way to combat this is to *ignore* all observations that contain `NA` values. Most statistics functions (like `mean()`) have an optional argument `na.rm`, which if set to `TRUE`, will ignore `NA`s when performing the calculation:

```{r}
data_missing %>%
  summarize(mean_x = mean(x, na.rm = TRUE))
```