---
title: "2.5 — OLS: Precision and Diagnostics — R Practice"
author: "YOUR NAME"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
    toc: true 
    toc_depth: 3
    toc_float: true
    code_folding: show
    highlight: tango
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

How did the results of the 2016 election in their district affect how often Members of Congress during the Trump Administration voted with the President? 

First, as always, load `tidyverse`!

## Question 1
Our [data](https://github.com/fivethirtyeight/data/tree/master/congress-trump-score) comes from fivethirtyeight's [Trump Congress tracker](https://projects.fivethirtyeight.com/congress-trump-score/). Fivethirtyeight has a great [public archive](https://data.fivethirtyeight.com/) of all of the data they use for their articles. I have saved a copy to this website so you can download it below. Download and import (`read_csv()`) the data into an object.

- [<i class="fas fa-table"></i> `congress-trump-score.csv`](/data/congress-trump-score.csv)

The variables that we care about are

| Variable | Description |
|----------|-------------|
| `congress` | Congressional session |
| `chamber` | Chamber of Congress |
| `last_name` | Member of Congress |
| `party` | Political Party |
| `state` | U.S. State |
| `district` | Congressional district |
| `agree_pct` | Proportion of votes that agree with President Trump (0-1) |
| `net_trump_vote` | District's margin of victory in 2016 election (positive is Trump win, negative is Clinton win) |

---

<!--ANSWER BELOW HERE-->

```{r}
# PUT CODE HERE
```

---

## Question 2
Look at the data with `glimpse()`. How many variables are there? How many observations?

---

<!--ANSWER BELOW HERE-->

```{r}
# PUT CODE HERE
```

---

## Question 3
`agree_pct` is oddly named, given that its values range from 0 to 1. Make a new variable (you can overwrite `agree_pct`) that is a true percentage, from 0 to 100, and use this going forward (it will make interpretation of our results easier).

---

<!--ANSWER BELOW HERE-->

```{r}
# PUT CODE HERE
```

---

## Question 4
Make a scatterplot of `agree_pct` on `net_trump_vote`. Add a regression line by adding an additional layer of `geom_smooth(method = "lm")`.

---

<!--ANSWER BELOW HERE-->

```{r}
# PUT CODE HERE
```

---

## Question 5
Find the correlation between `agree_pct` and `net_trump_vote`. Hint: if using `tidyverse`, like calculating any statistic, we want the `summarize()` the data. 

---

<!--ANSWER BELOW HERE-->

```{r}
# PUT CODE HERE
```

---

# Regression

## Question 6
We want to predict the following model:

$$\widehat{\text{agree_pct}}= \hat{\beta_0}+\hat{\beta_1} \,\text{net_trump_vote}$$
Run a regression and save it as an object. Then get a `summary()` of the object.

---

<!--ANSWER BELOW HERE-->

```{r}
# PUT CODE HERE
```

---
 
### Part A
What is $\hat{\beta_0}$ for this model? What does it mean in the context of our question?

---

<!--ANSWER BELOW HERE-->

```{r}
# PUT CODE HERE
```

---

### Part B
What is $\hat{\beta_1}$ for this model? What does it mean in the context of our question?

---

<!--ANSWER BELOW HERE-->

```{r}
# PUT CODE HERE
```

---

### Part C
What is $R^2$ for this model? What does it mean in the context of our question?

---

<!--ANSWER BELOW HERE-->

```{r}
# PUT CODE HERE
```

---

### Part D
What is the $SER$ for this model? What does it mean in the context of our question?

---

<!--ANSWER BELOW HERE-->

```{r}
# PUT CODE HERE
```

---

## Question 7
We can look at regression outputs in a tidier way, with the `broom` package. 

### Part A
Install (if you have not yet done so) and load the `broom` package.

---

<!--ANSWER BELOW HERE-->

```{r}
# PUT CODE HERE
```

---

### Part B
Run the function `tidy()` on your regression object (saved in question 6). Save this result as an object and then look at it.

---

<!--ANSWER BELOW HERE-->

```{r}
# PUT CODE HERE
```

---

### Part C
Run the `glance()` function on your original regression object. What does it show you? Find $R^2$ and the SER and confirm they are the same from the Base R `lm` output.

---

<!--ANSWER BELOW HERE-->

```{r}
# PUT CODE HERE
```

---

### Part D
Now run the `augment()` function on your original regression object, and save this as an object. Look at it. What does it show you?

---

<!--ANSWER BELOW HERE-->

```{r}
# PUT CODE HERE
```

---

## Question 8

Let's use our `broom` results to calculate the goodness of fit statistics you found in question 6 to confirm.

Calculate $R^2$ as $\frac{ESS}{TSS}$ by taking the variance of $\widehat{\text{agree_pct}_i}$ (`.fitted` in the `augment`ed object you made in Question 6D) over the variance of $\text{agree_pct}$.

Alternately, you can calculate $R^2$ as $1-\frac{SSE}{TSS}$ by taking 1 minus sum of squared $\hat{u_i}$ (`.resid` in that same `augment`ed object) over the variance of $\text{agree_pct}$.

**Optional**: If you really want to be fancy, make your own function to calculate the sum of squares of $\hat{Y}$ and $Y_i$ (instead of variance), as I did, with:

```{r, eval = F}
sum_sq = function(x){sum((x - mean(x))^2)}
```

and then running this function on `agree_pct` and `.fitted`.

---

<!--ANSWER BELOW HERE-->

```{r}
# PUT CODE HERE
```

---

## Question 9

Now let's try presenting your results in a regression table. Install and load the `huxtable` package, and run the `huxreg()` command. Your main input is your regression object you saved in Question 6. Feel free to customize the output of this table (see the slides).

---

<!--ANSWER BELOW HERE-->

```{r}
# PUT CODE HERE
```

---

# Regression Diagnostics

## Question 10
Now let's start looking at the residuals of the regression. 

### Part A
Take the `augment`ed regression object from Question 7D and use it as the source of your data to create a histogram with `ggplot()`, setting `aes(x = .resid)`. Does it look roughly normal?

---

<!--ANSWER BELOW HERE-->

```{r}
# PUT CODE HERE
```

---

### Part B

Take the `augment`ed regression object and make a residual plot, which is a scatterplot where `x` is the normal `x` variable, and `y` is the `.resid`. Feel free to add a horizontal line at 0 with `geom_hline(yintercept = 0)` as an additional layer. 

---

<!--ANSWER BELOW HERE-->

```{r}
# PUT CODE HERE
```

---

## Question 11
Now let's check for heteroskedasticity.

### Part A

Looking at the scatterplot and residual plots in Questions 3 and 7B, do you think the errors are heteroskedastic or homoskedastic?

---

<!--ANSWER BELOW HERE-->

```{r}
# PUT CODE HERE
```

---

### Part B

Install and load the `lmtest` package and run `bptest()` on your saved `lm` object from Question 6. According to the test, is the data heteroskedastic or homoskedastic?

---

<!--ANSWER BELOW HERE-->

```{r}
# PUT CODE HERE
```

---

### Part C

Now let's make some heteroskedasticity-robust standard errors. Install and load the `estimatr` package and use the `lm_robust()` command (instead of the `lm()` command) to run a new regression (and save it). Make sure you add `se_type = "stata"` inside the command to calculate robust SEs (in the same way that the Stata software does...long story). Look at it. What changes?

---

<!--ANSWER BELOW HERE-->

```{r}
# PUT CODE HERE
```

---

### Part D

Now let's see this in a nice regression table. Use `huxreg()` again, but add both your original regression and your regression saved in part C. Notice any changes?

---

<!--ANSWER BELOW HERE-->

```{r}
# PUT CODE HERE
```

---

## Question 12

Now let's check for outliers. 

### Part A

Just looking at the scatterplot in Question 3, do you see any outliers?

---

<!--ANSWER BELOW HERE-->

```{r}
# PUT CODE HERE
```

---

### Part B

Install and load the `car` package. Run the `outlierTest` command on your regression object. Does it detect any outliers?

---

<!--ANSWER BELOW HERE-->

```{r}
# PUT CODE HERE
```

---

### Part C

Look in your original data to match this outlier with an observation. Hint: use the `slice()` command, as the outlier test gave you an observation (row) number!

---

<!--ANSWER BELOW HERE-->

```{r}
# PUT CODE HERE
```

---

## Question 13 

(**Optional**: Flexing your tidyverse skills)

This data is still a bit messy. Let's check your `tidyverse` skills again! For example, we'd probably like to plot our scatterplots with colors for Republican and Democratic party. Or plot by the House and the Senate.

### Part A

First, the variable `congress` (session of Congress) seems a bit off. Get a `count()` of `congress`.

---

<!--ANSWER BELOW HERE-->

```{r}
# PUT CODE HERE
```

---

### Part B

Let's get rid of the `0` values for `congress` (someone made a mistake coding this, probably).

---

<!--ANSWER BELOW HERE-->

```{r}
# PUT CODE HERE
```

---

### Part C

The variable `party` is also quite a mess. `count()` by `party` to see. Then let's `mutate` a variable to make a better measure of political party - just `"Republican"`, `"Democrat"`, and `"Independent"`. Try doing this with the `case_when()` command (as your `mutate` formula).

The syntax for `case_when()` is to have a series of `condition ~ "Outcome"`, separated by commas. For example, one condition is to assign both `"Democrat"` and `"D"` to `"Democrat"`, as in `party %in% c("Democrat", "D") ~ "Democrat"`. You could also do this with a few `ifelse()` commands, but that's a bit more awkward.] When you're done `count()` by your new party variable to make sure it worked.

---

<!--ANSWER BELOW HERE-->

```{r}
# PUT CODE HERE
```

---

### Part D
Now plot a scatterplot (same as Question 4) and set `color` to your party variable. Notice `R` uses its own default colors, which don't match to the actual colors these political parties use! Make a vector where you define the party colors as follows: 

```{r, eval=F}
party_colors <- c("Democrat" = "blue",
                  "Republican" = "red",
                  "Independent" = "gray")
```

Then, run your plot again, adding the following layer to customize the colors `scale_colour_manual("Parties", values = party_colors)`. `"Parties"` is the title that will show up on the legend, feel free to edit it, or remove the legend with another layer `+guides(color = F)`.

---

<!--ANSWER BELOW HERE-->

```{r}
# PUT CODE HERE
```

---

### Part E

Now facet your scatterplot by `chamber` by adding a layer: `facet_wrap(~chamber)`.

---

<!--ANSWER BELOW HERE-->

```{r}
# PUT CODE HERE
```

---
