---
title: "Class Size & Test Scores Regression Analysis"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load-libraries, echo = F, warning = F, message = F}
# note I have already installed all of these here!

library(tidyverse) # always! 
library(haven) # for importing .dta files
library(broom) # for tidy regression commands
library(equatiomatic) # if you want to write reg. eq'ns in markdown
library(ggthemes) # for more ggplot themes
library(huxtable) # for regression output tables
library(lmtest) # for heteroskedasticity test
library(estimatr) # for robust-SEs
library(car) # for outlier test
```

# Loading In Data

```{r load-data}
CASchool <- read_dta("caschool.dta") # this is in the same directory as this Rmd file
```

```{r look-at-data-1}
# see what it gives us
# CASchool # don't run this in knitted markdown, it will print 420 rows!!
 
# note since we have tidyverse loaded, it prints a nice tibble

# look only at first 10 rows
head(CASchool, n = 10)

# get structure
str(CASchool)

# use tidyverse methods to look at structure
glimpse(CASchool)
```

# Scatterplot

```{r scatterplot, fig.retina=3}
# save as scatter
scatter <- ggplot(data = CASchool)+
  aes(x = str,
      y = testscr)+
  geom_point(color = "blue")+
  labs(x = "Student to Teacher Ratio",
       y = "Test Score")+
  theme_pander(base_family = "Fira Sans Condensed",
           base_size = 20)

# look at it 
scatter
```

# Simple Linear Regression (1 Variable)

```{r run-regression}
# run simple regression, save as "reg"
reg <- lm(testscr ~ str, data = CASchool)

# print it (just gives us the coefficients)
reg

# get full summary
summary(reg)

# Note: this also works (tidyverse): 
# reg %>% summary()
```

```{r broom-regression}
# use broom for tidier output

# tidy() gives us coefficients
reg %>% tidy()

# glance() gives us regression fit statistics
reg %>% glance()

# augment() creates observations for fitted and residual values (among other things)
reg %>% augment() %>% head() # I added head() because otherwise it will print 420 rows!!
```

You can type out the regression equation in markdown using LaTeX math code:

$$\widehat{\text{test score}}=689.9-2.28 \times str$$ 

Or you can try out `equatiomatic`:

```{r, results="asis"}
# be sure to set chunk option to: results="asis"

extract_eq(reg, # regression lm object
           use_coefs = TRUE, # use names of variables
           coef_digits = 2, # round to 2 digits
           fix_signs = TRUE) # fix negatives (instead of + -)
```

Now let's go back to the scatterplot and add our regression line!

```{r scatter-with-regression}
scatter+
  geom_smooth(method = "lm", # at minimum, must set method to linear model
              color = "red", # change color
              se = TRUE) # set to FALSE if you want to hide the SE range (gray)
```

Now let's print a nice regression output table

```{r basic-output-table}
huxreg(reg)
```

We can customize it a bit

```{r fancier-output-table}
huxreg("Test Score" = reg, # set Y name
       coefs = c("Intercept" = "(Intercept)", # rename X variable & intercept
                 "STR" = "str"),
       statistics = c("N" = "nobs", # show only these stats and name them
                      "R-Squared" = "r.squared",
                      "SER" = "sigma"),
       number_format = 2) # round to 2 decimals
```


# Regression Analysis

```{r, residuals_histogram}
# first let's use augment again, and *save* as aug_reg
aug_reg <- augment(reg)

# now we can plot a histogram using aug_reg

ggplot(data = aug_reg)+
  aes(x = .resid)+ # the residuals!
  geom_histogram(color = "white",
                 fill = "hotpink")+
  labs(x = expression(paste("Residual, ", hat(u))))+
  theme_pander(base_family = "Fira Sans Condensed",
           base_size=20)
```

```{r, residual-plot}
ggplot(data = aug_reg)+
  aes(x = str,
      y = .resid)+
  geom_point(color="blue")+
  geom_hline(aes(yintercept = 0), color="red")+
  labs(x = "Student to Teacher Ratio",
       y = expression(paste("Residual, ", hat(u))))+
  theme_pander(base_family = "Fira Sans Condensed",
           base_size=20)
```

Testing for heteroskedasticity

```{r check-heteroskedasticity}
# using package lmtest
bptest(reg)
```

```{r robust-SE-regression}
# using package estimatr
reg_robust <-lm_robust(testscr ~ str,
                              data = CASchool,
                              se_type = "stata")

reg_robust %>% summary()
```

```{r reg-output-robust}
huxreg("Normal" = reg,
       "Robust" = reg_robust,
       coefs = c("Intercept" = "(Intercept)",
                 "STR" = "str"),
       statistics = c("N" = "nobs",
                      "R-Squared" = "r.squared",
                      "SER" = "sigma"),
       number_format = 2)
```

Check for outliers

```{r test-outliers}
# using car package
outlierTest(reg) # will point out which obs #s seem outliers
```

# Multivariate Regressions

## Correlation Table

```{r}
CASchool %>%
  select("str","testscr","el_pct", "expn_stu") %>%
  cor()
```

## Conditional Distribution (of Test Score) by Low/High ESL

```{r}
# make a new variable called EL
# = high (if el_pct is above median) or = low (if below median)
CASchool<-CASchool %>% # next we create a new dummy variable called ESL
  mutate(ESL = ifelse(el_pct > median(el_pct), # test if ESL is above median
                     yes = "High ESL", # if yes, call this variable "High ESL"
                     no = "Low ESL")) # if no, call this variable "Low ESL"

# get average test score by high/low EL
CASchool %>%
  group_by(ESL) %>%
  summarize(Average_test_score=mean(testscr))

# make distributions
ggplot(data = CASchool)+
  aes(x = testscr,
      fill = ESL)+
  geom_density(alpha=0.5)+
  labs(x = "Test Score",
       y = "Density")+
  ggthemes::theme_pander(
    base_family = "Fira Sans Condensed",
    base_size=20
    )+
  theme(legend.position = "bottom")

# make scatterplot

esl_scatter<-ggplot(data = CASchool)+
  aes(x = str,
      y = testscr,
      color = ESL)+
  geom_point()+
  geom_smooth(method="lm")+
  labs(x = "STR",
       y = "Test Score")+
  ggthemes::theme_pander(
    base_family = "Fira Sans Condensed",
    base_size=20
    )+
  theme(legend.position = "bottom")
esl_scatter

# facet instead

esl_scatter+
  facet_grid(~ESL)+ #<<
  guides(color = F) #<<
```

## Multivariate Regression (with %EL)

```{r}
# run regression of testscr on str and el_pct
school_reg_2 <- lm(testscr ~ str + el_pct, 
                 data = CASchool)

# look at it (just gives coefficients)
school_reg_2

# get summary
school_reg_2 %>% summary()
```

```{r}
# do it tider with broom
school_reg_2 %>% tidy()

# get regression fit statistics
school_reg_2 %>% glance()
```

```{r}
# make a regression output table
library(huxtable)
huxreg("Model 1" = reg,
       "Model 2" = school_reg_2,
       coefs = c("Intercept" = "(Intercept)",
                 "Class Size" = "str",
                 "%ESL Students" = "el_pct"),
       statistics = c("N" = "nobs",
                      "R-Squared" = "r.squared",
                      "SER" = "sigma"),
       number_format = 2)
```

## Measuring OVB

```{r}
# "true" regression (str and %EL)
true<-lm(testscr~str+el_pct, data=CASchool)
true %>% tidy()

# "omitted" regression (omitting %EL)
omitted<-lm(testscr~str, data=CASchool)
omitted %>% tidy()

# "auxiliary" regression (%EL on str)
aux<-lm(el_pct~str, data=CASchool)
aux %>% tidy()
```

# Variance and VIF

```{r}
library(car) # needs car package
true %>% vif()


# calculate VIF manually

# look at aux reg stats for R^2
# extract our R-squared from aux regression (R_j^2)
aux_r_sq<-glance(aux) %>%
  select(r.squared)

aux_r_sq # look at it

# calculate VIF manually

our_vif<-1/(1-aux_r_sq) # VIF formula 

our_vif

```

